%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------


\documentclass[preprint]{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{azalea}


\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country} 
\copyrightyear{20yy} 
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
\doi{nnnnnnn.nnnnnnn}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

%\titlebanner{banner above paper title}        % These are ignored unless
%\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{CoLoSL: Concurrent Local Subjective Logic}
\subtitle{}

\authorinfo{Azalea Raad\and Philippa Gardner\and Jules Villard}
           {Imperial College London}
           {\{azalea, pg, j.villard\}@imperial.ac.uk}
%\authorinfo{Name2\and Name3}
%           {Affiliation2/3}
%           {Email2/3}

\maketitle

\begin{abstract}
  A key difficulty in verifying the correctness of programs operating
  concurrently over a shared memory is being able to reason about each
  component in isolation of the others, even though in reality the
  correctness of the whole system is the collaborative result of
  intricately intertwined actions of all components.  Compositionality
  of the reasoning is crucial if we are to ever verify large
  concurrent systems, and also to replicate programmers intuitions
  about why their implementations are correct. In the sequential
  world, separation logic has demonstrated that a key ingredient to
  achieve compositional reasoning is the ability to reason only about
  portions of the machine state actually accessed by programs (their
  footprints). Unfortunately, existing techniques for shared-memory
  concurrency verification force us to reason globally about the
  shared state and interferences, regardless of which portion of it is
  actually used by the component under scrutiny.

  This paper introduces the program logic \colosl, in which threads
  can be verified with respect to their \emph{subjective view} of the
  shared state, which may overlap with other threads' subjective
  views in arbitrary ways. Interferences by the environment can be
  rewritten as long as the projected effect on the subjective view
  remains the same, and interferences on parts of the shared state
  outside of the current view can be safely forgotten. This
  flexibility yields compositional proofs for shared-memory
  concurrency, which we demonstrate on a range of examples including a
  concurrent computation of the spanning tree of a graph.
\end{abstract}

\begin{abstract}

\pgcomment{I've used thread instead of component, just pick one word
  to describe one thing.}

  A key difficulty in verifying shared-memory concurrent programs
is reasoning 
  compositionally  about each
 thread  in isolation, even though in reality the
  correctness of the whole system is the collaborative result of
  intricately intertwined actions of the threads.  Such compositionality
  is essential for verifying  large
  concurrent systems and replicating a  programmer's intuition
  about why their implementations are correct. For sequential
programs, separation logic has demonstrated the importance of reasoning 
only about
  portions of the machine state (the resource) actually accessed by
  programs. For concurrent programs,  existing techniques 
require global reasoning about shared resource, even though a
thread might only access a small part of it.


  This paper introduces the program logic \colosl, in which threads
  are verified with respect to their \emph{subjective views}. Each
  subjective (personalised) view provides a thread-specific
  description of the shared state, comprising the partial shared
  resource necessary for the thread to run and a thread-specific {
    interference relation} describing how the thread and the
  environment can effect this partial shared resource. Subjective
  views may arbitrarily overlap with each other.  In particular, the
  interference relations expand and contract with relation to the
  partial shared resource, as long as the projected effect on the
  subjective view remains the same.  This flexibility provides truely  
  compositional proofs for shared-memory concurrency, which we
  demonstrate on a range of examples including a concurrent
  computation of the spanning tree of a graph.
\end{abstract}


\category{CR-number}{subcategory}{third-level}

% general terms are not compulsory anymore, 
% you may leave them out
%% \terms
%% term1, term2

\keywords
Concurrency, Program logic, Compositional reasoning, Separation logic

\input{Sections/Introduction/introduction.tex}
\input{Sections/Intuition/intuition.tex}
\input{Sections/Logic/logic.tex}
\input{Sections/SemanticsAndSoundness/semantics.tex}
\input{Sections/FurtherExamples/examples.tex}
\input{Sections/Conclusions/conclusions.tex}

% POPL recommends abbrvnat bibliography style.
%% \bibliographystyle{abbrvnat}
\azaleacomment{Need to anonymise the tech report.}
\bibliographystyle{plain_better}
\bibliography{biblio}
% The bibliography should be embedded for final submission.


\end{document}
